---
title: "Practical Machine Learning Assignment"
author: "Chun-Fu Wang"
output: html_document
---

# Loading the Data
```{r}
library(dplyr)
library(caret)

trainData <- read.csv("pml-training.csv")
testData <- read.csv("pml-testing.csv")

trainData <- trainData[2:length(trainData)]
testData <- testData[2:length(testData)]

dim(trainData)
dim(testData)

```
There are `{r} I(length(trainData))` observations in the training data set and `{r} I(length(testData))` observations in the test data set that we are going to predict.


# Exploratory Data Analysis
Removing the near zero variance features as well as the statistically insignificant features.

```{r}
# cleaning up data
dim(trainData)
nzv <- nearZeroVar(trainData)
filteredTrainData <- trainData[, -nzv]
filteredTestData <- testData[, -nzv]

# removed statistically insignificant variables
filteredTrainData <- 
    filteredTrainData %>% 
    select(-c(user_name, 
              raw_timestamp_part_1, 
              raw_timestamp_part_2, 
              cvtd_timestamp,
              max_roll_belt:var_yaw_belt,
              var_accel_arm,
              max_picth_arm:amplitude_yaw_arm,
              max_roll_dumbbell:amplitude_pitch_dumbbell,
              var_accel_dumbbell:var_yaw_dumbbell,
              max_picth_forearm:amplitude_pitch_forearm,
              var_accel_forearm))

# remove from test set as well
filteredTestData <- 
    filteredTestData %>% 
    select(-c(user_name, 
              raw_timestamp_part_1, 
              raw_timestamp_part_2, 
              cvtd_timestamp,
              max_roll_belt:var_yaw_belt,
              var_accel_arm,
              max_picth_arm:amplitude_yaw_arm,
              max_roll_dumbbell:amplitude_pitch_dumbbell,
              var_accel_dumbbell:var_yaw_dumbbell,
              max_picth_forearm:amplitude_pitch_forearm,
              var_accel_forearm))

dim(filteredTrainData)
dim(filteredTestData)

```

# Preprocess
Split the training data into two set of 80% and 20%.
```{r}
set.seed(142678)

dataIndex <- createDataPartition(filteredTrainData$classe, p = 0.8, list = FALSE)
trainSet <- filteredTrainData[dataIndex, ]
testSet <- filteredTrainData[-dataIndex, ]

```

# Machine Learning
Using Random Forest and Rpart to train the model.
```{r}
library(doParallel)
registerDoParallel()

modelRf <- train(classe ~ ., data = trainSet, model = "rf")
modelRpart <- train(classe ~ ., data = trainSet, model = "rpart")

predRf <- predict(modelRf, newdata = testSet)
predRpart <- predict(modelRpart, newdata = testSet)

C1 <- confusionMatrix(predRf, testSet$classe)
print(C1)

C2 <- confusionMatrix(predRpart, testSet$classe)
print(C2)

```

# Out of Sample Error
Out of sample error for Random Forest is: `{r} I((1 - C1$overall[1]) * 100)`.
Out of sample error for Decision Tree is: `{r} I((1 - C2$overall[1]) * 100)`.

# Predictin Result
Predicting the result using Random Forest models because of the lower out of sample error.
```{r}
predResult <- predict(modelRf, newdata = filteredTestData)
print(predResult)

```
